Online Classification of arXiv Papers Based on Your Preferences

### 2014-05-14 Paper.tex
paper updates

\documentclass[12pt]{article}
\usepackage{amsmath,amsthm,amssymb}
\usepackage{graphicx}
\pagestyle{myheadings}
\markright{Joseph McDonald}

@@

One of the important aspects of the algorithm is that it present some articles to the user that it predicted to be uninteresting. Otherwise there is risk of mislabeling and potentially discarding many articles that are interesting to the user. Nevertheless the goal is to have an active filter on the articles lest the whole purpose of the program be defeated. Since there must be a balance between presenting only interesting articles and having a full understanding of the user's interests, we must choose what fraction of uninteresting articles are presented. In the following section we present a bound on the number of updates made by the Perceptron algorithm we employ, justifying why the algorithm may work towards an accurate hypothesis while receiving labels for only a fraction of the samples labeled uninteresting.

@@

Let $I_1$ be the set of $t$ where a mistake was made and
$\widehat{y}=1$ (false positives).  Let $I_2$ be the set where a mistake was made with
$\widehat{y} = -1$ and $k = 1$ (false negatives where the algorithm updates), and $I_3$ the set where a mistake was
made with $\widehat{y} = -1$ and $k = 0$ (false negatives where the algorithm does {\it not} make an update). Finally, let $M_i = |I_i|$ for $i =
1,2,3$. We bound $M_1 + M_2$ in the same way we bound $M$ in the original Perceptron algorithm. Assume there exists a $\rho$ and $R$ satisfying the assumptions of Theorem 7.8 from our text, and a separating hyperplane determined by the vector $v$. Let $M' = M_1 + M_2$, then
\begin{align*}
M' \rho =&\ \frac{v}{\|v\|}\sum_{I_1 \cup I_2} y_tx_t \
 \leq   \Bigl\|\sum_{t \in I_1 \cup I_2} y_t x_t\Bigr\| \\
=& \ \Bigl\|\sum_{t \in I_1 \cup I_2} (w_{t+1} - w_t) \Bigr\| 
= \ \|w_{T+1} \| \\
=&\ \sqrt{\sum_{t \in I_1 \cup I_2} \|  w_{t+1}\|^2 - \|w_t\|^2 }\\
= &\ \sqrt{\sum_{t \in I_1 \cup I_2} \|  w_{t} + y_t x_t\|^2 - \|w_t\|^2 }\\
= &\ \sqrt{\sum_{t \in I_1 \cup I_2} 2y_t w_t\cdot x_t - \|x_t\|^2} \\
\leq &\ \sqrt{\sum_{t \in I_1 \cup I_2} \|x_t\|^2}
\leq \sqrt{M'r^2}.
\end{align*}
This implies that $M' \leq \frac{r^2}{\rho^2}$. Since $M = M' + M_3$, a bound on $M_3$ is needed. We revert to a weaker bound and consider the expectation of
$M_3$. Since 
\[M_3 = \sum_{t=1}^T 1_{\left\{\widehat{y}_t = -1 \cap y_t = 1\right\}}1_{k_t = 0}, \]
we get that
\begin{align*}
\E[M_3] = & \sum_{t=1}^T \E[1_{\left\{\widehat{y_t} = -1 \cap y_t = 1\right\}}]\E[1_{k_t = 0}] \\
 = &\ (1 - p)\sum_{t=1}^T \E[1_{\left\{\widehat{y_t} = -1 \cap y_t = 1\right\}}],
\end{align*}
where the last equality uses the independence of $k_t$ from the data. To control the sum of expectations in this expression we consider the expectation of $M_2$. Since
\[
M_2 =  \sum_{t=1}^T 1_{\left\{\hat{y_t} = -1 \cap y_t = 1\right\}}1_{k_t = 1},
\]
then
\begin{align*}
\E[M_2] =&\ \sum_{t=1}^T \E[1_{\left\{\hat{y_t} = -1 \cap y_t = 1\right\}}]\E[1_{k_t = 1}] \\
=&\ p\sum_{t=1}^T \E[1_{\left\{\hat{y_t} = -1 \cap y_t = 1\right\}}].
\end{align*}
Thus
\[
E[M_3] = \ \frac{1-p}{p}\ \E[M_2].
\]
We can apply the crude bound $\E[M_3] \leq \frac{1-p}{p}\ \E[M']$, since $M_2 \leq M'$. Thus the expected number of mistakes can be bounded by
\begin{align*}
\E[M]  =  &\ \E[M'] + \E[M_3] \\
\leq & \ \frac{r^2}{\rho^2} + \frac{1-p}{p}\frac{r^2}{\rho^2} \\
\leq  & \ \frac{1}{p}\frac{r^2}{\rho^2}
\end{align*}
Note that this is a weak bound.  Not only is it greater than the
original perceptron bound by a factor of $\frac{1}{p}$, it is also only a
bound in expectation.\\

This algorithm can be also be used with other classifiers.  The set of data on which the scheme updates on is no longer iid, but any bound on the number of mistakes that doesn't require iid data, like the bound for Perceptron algorithm, will still hold in expectation with a similar modification. We apply the same approach to the OnlineSVM, Margin Perceptron, and Kernel Perceptron algorithms.  \\
The general method outlined above corrects negative predictions slowly and at a constant rate.  It seems natural that instead we might want to correct a lot of negative predictions when we feel the algorithm is not accurate, and be more lenient when the algorithm is doing well.  In order to achieve this, we introduce an adaptive $p$ scheme.  This scheme starts with some value of $p$, and each time it asks the user to label a data point that it predicted would have a negative label, it modifies $p$.  If it was correct, then $p$ is reduced and it checks negative predictions less often.  On the other hand, if it was incorrect it increases $p$. \\

\noindent Adaptive $p$ Perceptron algorithm:
\begin{quotation}
%\noindent Modified Perceptron:\\
\noindent $w_1=0$\\
\noindent $p=1$\\
{\tt for $t=1$ to $T$}:\\
\indent receive($x_t$)\\
\indent $\widehat{y}_t=\mbox{sign}(w_t\cdot x_t)$\\
\indent {\tt if} $\widehat{y}_t =1$ : \\
\indent\indent receive($y_t$)\\
\indent\indent {\tt if $y_t\neq\widehat{y}_t$}:\\
\indent\indent\indent $w_{t+1} = w_t + y_t x_t$\\
\indent\indent {\tt else $w_{t+1} = w_t$}\\
\indent\{\tt else if ($\widehat{y}_t =-1$ {\tt and} $k_t = 1$):\\
\indent\indent {\tt if $y_t\neq\widehat{y}_t$}:\\
\indent\indent\indent $w_{t+1} = w_t + y_t x_t$\\
\indent\indent\indent $p=p + \gamma_1(1 - p)$\\
\indent\indent {\tt else }
\indent\indent\indent $w_{t+1} = w_t + y_t x_t$\\
\indent\indent\indent $p = p - \gamma_2(p - p_{\min})$}
\indent {\tt else if $\widehat{y}_t = -1$ and $k_t=0$}:\\
\indent\indent $w_{t+1} = w_t$\\
{\tt return $w_{T+1}$}
\end{quotation}
Where $p_{\min}$, $\gamma_1$ and $\gamma_2$ are parameters that can be tuned by the user. It is easy to extend the bound above to this case.  Let $I_i$, $M_i$, and $M'$ be defined as above. The bound $M'\leq r^2/\rho^2$ follows in exactly the same way.  We then bound $M_3$ in expectation, 
\begin{align*}
\E[M_3] = & \sum_{t=1}^T \E[1_{\left\{\widehat{y_t} = -1 \cap y_t = 1\right\}}]\E[1_{k_t = 0}] \\
 \leq &\ (1 - p_{\min})\sum_{t=1}^T \E[1_{\left\{\widehat{y_t} = -1 \cap y_t = 1\right\}}],
\end{align*}
and bound $\E[M_2]$ from below
\begin{align*}
\E[M_2] =&\ \sum_{t=1}^T \E[1_{\left\{\hat{y_t} = -1 \cap y_t = 1\right\}}]\E[1_{k_t = 1}] \\
\geq &\ p\sum_{t=1}^T \E[1_{\left\{\hat{y_t} = -1 \cap y_t = 1\right\}}].
\end{align*}
Continuing as before, this gives the following bound on $\E[M]$,
\begin{align}
\E[M] = &\E[M'] + \E[M_3] \leq \frac{r^2}{\rho^2} + \frac{1-p_{\min}}{p_{\min}} \frac{r^2}{\rho^2}\\
= & \frac{1}{p_{\min}}\frac{r^2}{\rho^2}\\
\end{align*}
Unfortunately, we do not have a stonger theoretical bound for this adaptive $p$ approach, but we test this scheme in section \ref{sec:results} and see that it performs fairly well while requiring less input from the user.\\

\section{Implementation}
The tool is made from components in multiple Python files, split up according to their function in the program.
The metadata for each article is obtained using arXiv's Open Archives Initiative protocol which returns an XML file with a list of up to 1000 articles as nodes containing a variety of metadata, including abstracts, authors, submission dates, and fields and subfields, making the data quite accessible and fairly simple to sort.
The file {\tt ArxivData.py} contains the functions and class definitions used to collect, organize and process data. Included are a class of objects representing papers and a class definition for feature vectors.
%
%DESCRIBE BAG OF WORDS APPROACH FOR FEATURE VECTORS SOMEWHERE
%
The primary function is {\tt GetPapersOAI()} and its wrapper which prompts the user for a particular subject, any of arXiv's listed subfields for that subject, and a range of dates from which to select the papers.
Then it draws the XML file of data from arXiv's database, filters the articles using the XML entries according to the subfields, and prepares the stream of abstracts to present the user.
The user is then asked to label each article as interesting or not, one at a time, with updates made according to the algorithm employed.

@@

An important feature of any text classification device is the data used to construct feature vectors. There are several approaches to text classification, itself a large field of research. We take an elementary approach and use a bag-of-words model, that is, performing a word count for each distinct word in the abstract and converting that into a vector. 
Each distinct word corresponds to a different coordinate in the feature space. This ``vectorization" is defined in {\tt ArxivData.py} and called when

%One of the most important aspects of this tool is what is used for feature data. There are several approaches to text categorization that 
\section{Results}
\label{sec:results}

We tested different online learning algorithms with sample data generated by graduate students at NYU.  Several students were asked to use a python script to gather the abstracts from Arxiv in categories that they specified (see SECTION REFERENCE HERE).  Then the students were asked to generate labels for each entry indicating whether they would be interested in reading a paper based on the abstract.  \\
This data was used to test the following online learning algorithms: Perceptron, Margin Perceptron, Online SVM, Gaussian Kernel Perceptron, and Polynomial Kernel Perceptron.  For each algorithm with tunable parameters, a grid search was performed to determine the best value of each parameter. For each combination of parameters, each classifier was separately trained on the data from each student, and the resulting performance was averaged over students.  Parameters chosen with a grid search are given in table \ref{table:Parametertable}.  \\

\begin{table}[h]
\centering
\begin{tabular}{l | r}
Algorithm & Parameters Tuned \\
\hline
Margin Perceptron &  Margin, $\rho$  \\
Online SVM & Penalty, $C$ \\
Gaussian Kernel Perceptron &  variance, $\sigma$, \\
Polynomial Kernel Perceptron &  intercept, $c$, and degree, $d$ \\
\hline
\end{tabular}
\label{table:Parametertable}
\caption{Parameters tuned for each classifier used.}
\end{table}

For this particular classification problem, it turned out that overall accuracy was a poor indicator of performance.  Students tended to only label a small subset of papers as ``interesting'', so an algorithm which always predicted a label of $\hat{y} = -1$ (the paper was not interesting to the user) could achieve very high values of accuracy.  Instead, we looked at Recall and Precision.  Recall, $R$, is the percentage of papers labeled $y=1$ by the user that were successfully categorized by the algorithm, $\hat{y} = 1$.  Precision, $P$, is the percentage of papers labeled $\hat{y} = 1$ by the algorithm that were actually interesting to the user, $y = 1$.  The mathematical definitions are as follows 
\begin{align*}
R(h) = & \frac{\sum_{i=1}^m 1_{\left{h(x_i) = 1 \cap y_i =1\right}}{\sum_{i=1}^m 1_{\left{y_i = 1\right}} \\
P(h) = & \frac{\sum_{i=1}^m 1_{\left{h(x_i) = 1 \cap y_i =1\right}}}{{\sum_{i=1}^m 1_{\left{h(x_i) = 1\right}}. \\
\end{align*}
For each classifier and combination of parameters, the algorithm was shown 50 papers to obtain reasonable weights, and then as it continued to run, the number of correct positive labels, incorrect positive labels, and incorrect negative labels were recorded and used to calculate $R$ and $P$.  These values were then averaged over all students separately. The best average $R$ (over all parameters tested) and the best average $P$ are shown in figure \ref{fig:PrecisionandRecall}, along with the number of negative labels asked of the user.  We can see that the scheme with adaptive $p$ is able to achieve performance close to the the original algorithms but requiring less than half as many negative updates.

\begin{figure}
\includegraphics[width=0.7\textwidth]{EveryoneRecallPaper.pdf}
\includegraphics[width=0.7\textwidth]{EveryonePrecisionPaper.pdf}
\label{fig:PrecisionandRecall}
\caption{Performance of tested online classification algorithms using different values of $p$.  The adaptive $p$ algorithm performs relatively well compared to the original algorithm ($p = 1$), and requires less than half as many negative updates.  Majority vote classifier combines all others in this figure.}
\end{figure}

\section{Conclusion}

>>>>

\documentclass[12pt]{article}
\usepackage{amsmath,amsthm,amssymb}
\pagestyle{myheadings}
\markright{Joseph McDonald}

@@

One of the important aspects of the algorithm is that it present some articles predicted to be uninteresting. Otherwise there is risk in mislabeling and potentially discarding many articles that are interesting to the user. Nevertheless the goal is to have an active filter on the articles lest the whole purpose of the program be defeated. Since there must be a balance between presenting only interesting articles and having a full understanding of the user's interests, we must choose what fraction of uninteresting articles are presented. In the following section we present a bound on the number of updates made by the Perceptron algorithm we employ, justifying why the algorithm may work towards an accurate hypothesis while receiving labels for only a fraction of the samples labeled uninteresting.

@@

Let $I_1$ be the set of $t$ where a mistake was made and
$\widehat{y}=1$ (false positives).  Let $I_2$ be the set where a mistake was made with
$\widehat{y} = -1$ and $k = 1$ (false negatives where the algorithm updates), and $I_3$ the set where a mistake was
made with $\widehat{y} = -1$ and $k = 0$ (false negatives where the algorithm does {\it not} make an update). Finally, let $M_i = |I_i|$ for $i =
1,2,3$. We bound $M_1 + M_2$ in the same way we bound $M$ in the original Perceptron algorithm. Assume there exists a $\rho$ and $R$ satisfying the assumptions of Theorem 7.8 from our text, and a separating hyperplane determined by the vector $v$. Let $M' = M_1 + M_2$, then
\begin{align*}
M' \rho =&\ \frac{v}{\|v\|}\sum_{I_1 \cup I_2} y_tx_t \
 \leq   \Bigl\|\sum_{t \in I_1 \cup I_2} y_t x_t\Bigr\| \\
=& \ \Bigl\|\sum_{t \in I_1 \cup I_2} (w_{t+1} - w_t) \Bigr\| 
= \ \|w_{T+1} \| \\
=&\ \sqrt{\sum_{t \in I_1 \cup I_2} \|  w_{t+1}\|^2 - \|w_t\|^2 }\\
= &\ \sqrt{\sum_{t \in I_1 \cup I_2} \|  w_{t} + y_t x_t\|^2 - \|w_t\|^2 }\\
= &\ \sqrt{\sum_{t \in I_1 \cup I_2} 2y_t w_t\cdot x_t - \|x_t\|^2} \\
\leq &\ \sqrt{\sum_{t \in I_1 \cup I_2} \|x_t\|^2}
\leq \sqrt{M'r^2}.
\end{align*}
This implies that $M' \leq \frac{r^2}{\rho^2}$. Since $M = M' + M_3$, a bound on $M_3$ is needed. We revert to a weaker bound and consider the expectation of
$M_3$. Since 
\[M_3 = \sum_{t=1}^T 1_{\left\{\widehat{y}_t = -1 \cap y_t = 1\right\}}1_{k_t = 0}, \]
we get that
\begin{align*}
\E[M_3] = & \sum_{t=1}^T \E[1_{\left\{\widehat{y_t} = -1 \cap y_t = 1\right\}}]\E[1_{k_t = 0}] \\
 = &\ (1 - p)\sum_{t=1}^T \E[1_{\left\{\widehat{y_t} = -1 \cap y_t = 1\right\}}],
\end{align*}
where the last equality uses the independence of $k_t$ from the data. To control the sum of expectations in this expression we consider the expectation of $M_2$. Since
\[
M_2 =  \sum_{t=1}^T 1_{\left\{\hat{y_t} = -1 \cap y_t = 1\right\}}1_{k_t = 1},
\]
then
\begin{align*}
\E[M_2] =&\ \sum_{t=1}^T \E[1_{\left\{\hat{y_t} = -1 \cap y_t = 1\right\}}]\E[1_{k_t = 1}] \\
=&\ p\sum_{t=1}^T \E[1_{\left\{\hat{y_t} = -1 \cap y_t = 1\right\}}].
\end{align*}
Thus
\[
E[M_3] = \ \frac{1-p}{p}\ \E[M_2].
\]
We can apply the crude bound $\E[M_3] \leq \frac{1-p}{p}\ \E[M']$, since $M_2 \leq M'$. Thus the expected number of mistakes can be bounded by
\begin{align*}
\E[M]  =  &\ \E[M'] + \E[M_3] \\
\leq & \ \frac{r^2}{\rho^2} + \frac{1-p}{p}\frac{r^2}{\rho^2} \\
\leq  & \ \frac{1}{p}\frac{r^2}{\rho^2}
\end{align*}
Note that this is a weak bound.  Not only is it greater than the
original perceptron bound by a factor of $\frac{1}{p}$, it is also only a
bound in expectation.

\section{Implementation}

The tool is made from components in multiple Python files, split up according to their function in the program.
The metadata for each article is obtained using arXiv's Open Archives Initiative protocol which returns an XML file with a list of up to 1000 articles as nodes containing a variety of metadata, including abstracts, authors, submission dates, and fields and subfields, making the data quite accessible and fairly simple to sort.
The file {\tt ArxivData.py} contains the functions and class definitions used to collect, organize and process data. Included are a class of objects representing papers and a class definition for feature vectors.
%
%DESCRIBE BAG OF WORDS APPROACH FOR FEATURE VECTORS SOMEWHERE
%
The primary function is {\tt GetPapersOAI()} and its wrapper which prompts the user for a particular subject, any of arXiv's listed subfields for that subject, and a range of dates from which to select the papers.
Then it draws the XML file of data from arXiv's database, filters the articles using the XML entries according to the subfields, and prepares the stream of abstracts to present the user.
The user is then asked to label each article as interesting or not, one at a time, with updates made according to the algorithm employed.

@@

An important aspect of any text classification device is the data used to construct feature vectors. There are several approaches to text classification, itself a large field of research. We take an elementary approach and use a bag-of-words model; that is, performing a word count for each distinct word in the abstract and converting that into a vector. 
Each distinct word corresponds to a different coordinate in the feature space. This ``vectorization" is defined in a class {\tt TextFeatureVector} in {\tt ArxivData.py} and is called in the function {\tt FeaturizeAbstracts()}, which appears before running the algorithms, naturally. Word counts seems like a naive approach, yet the bag-of-words model is well used and often augmented with other methods. Moreover implementing such a model is not terribly complicated, which makes it a useful and attractive option.

%One of the most important aspects of this tool is what is used for feature data. There are several approaches to text categorization that
